Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2017-07-07 Response.tex

The discussion on the variational integrators has been reduced. 
Much of the background material on Lagrangian mechanics is removed and an additional paragraph is included to summarize the key ideas of variational integrators. 
The additional paragraph is shown below.

\begin{correction}
Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invaraint properties and symplecticity.
Variational integrators are constructed by discretizing Hamilton's principle rather than the continuous Euler-Lagrange equations.
As a result, integrators developed in this manner have the desirable properties that they are symplectic and momentum preserving.
In addition, they exhibit improved energy behavior over long integration periods.
A thorough discussion of variational integrators is provided in~\cite{west2004,marsden2001}.
We use this approach to construct a variational integrator for the PCRTBP with the inclusion of low-thrust propulsion.

Consider the autonomous mechanical system described by the Lagrangian, \( L(q, \dot{q} ) \), for the generalized coordinates, \( q\) and velocities \( \dot{q} \).
The integration of the continuous Lagrangian along a path, \( q (t) \), followed by the system must satisfy Hamilton's principle, which results in the well-known Euler-Lagrange equations~\cite{lanczos1970}.
In the discrete time scenario, the continuous path \( q(t), \dot{q}(t) \), is replaced by a finite difference approximation over a fixed time step, e.g. \( q_0, \frac{q_1 - q_0}{h} \), which converges to the true velocity as \( h \) tends towards zero.
In this manner, we define a discrete Lagrangian, \( L_d (q_0, q_1, h) \) which approximates the integral of the true Lagrangian over the time interval \( h \) between \( q_0 \) and \( q_1\)~\cite{marsden2001}.
\end{correction}

\item 
    \begin{itshape}
The second-order trapezoidal rule will give significantly poorer state vector accuracy with respect to a high-precision RK integrator, even if the state is guaranteed to lie on an energy surface.  Why didn't the authors consider going to a higher-order variational integrator, or even a Taylor Integrator?  It is very likely that the integration accuracy will introduce some errors in the accuracy of the partial derivatives and in the states.
\end{itshape}

We chose the trapezoidal rule as it provides a second order quadrature approximation.
In addition, it provides an explicit rule that eases the complexity in using an implicit solution method.

\item 
    \begin{itshape}
Section 3.2 - please cite a reference as to where the selected quadrature rules presented in Table 1 have been derived.
\end{itshape}

In the interest of conciseness and to avoid repeating material that has already been covered in the literature, the discussion on variational integrators has been substantially reduced.
As a result, this table has been removed from the manuscript.

@@

Figure 2 - x-axis has no label but I think it's nondimensional time, please indicate.  Label all axes even if they are nondimensional units.

I recommend removing appendix B.  Giving the equations and stating that the Matrix inversion is from Gauss-Jordan elimination is probably sufficient, unless there is some compelling algebraic manipulation that is very unique to the problem.

In line 14 (page 12), you define the state transition matrix over one orbital period, and then on line 18 you start referring to it as the monodromy matrix.  I understand what you're doing, but you may want to make the definition more explicit.  Some of this section can be substantially reduced and other source material could be referenced anyway.

\item 
    \begin{itshape}
Comparison like that of Figure 3.4 are very common in the well-known literature concerning variational integrators.  Moreover, the proper way to present such a comparison is with different delta-T's in the integration step, showing when the Runge-Kutta methods actually break down in terms of maintaining an accurate portrayal of the Energy/Hamiltonian conservation.

### 2017-07-07 Response.tex

\item 
    \begin{itshape}
Comparison like that of Figure 3.4 are very common in the well-known literature concerning variational integrators.  Moreover, the proper way to present such a comparison is with different delta-T's in the integration step, showing when the Runge-Kutta methods actually break down in terms of maintaining an accurate portrayal of the Energy/Hamiltonian conservation.

For example -

Figure 2.4
Lew, A., Marsden, J., Ortiz, M., West, M., "An Overview of Variational Integrators," FINITE ELEMENT METHODS: 1970's AND BEYOND L.P. Franca (Ed.) c CIMNE, Barcelona, Spain 2003.

Online copy:
http://lagrange.mechse.illinois.edu/pubs/LeMaOrWe2004/LeMaOrWe2004.pdf
\end{itshape}

The comparison plot has been modified to follow the precedent in the literature.
The plot shows the deviation of the Jacobi constant as the time step is varied. 
In addition, an additional plot has been added, and some related discussion which shows the decrease in computation cost through the use of the variational integrator.

### 2017-07-07 Response.tex

\item 
    \begin{itshape}
Section 3.2 - please cite a reference as to where the selected quadrature rules presented in Table 1 have been derived.
\end{itshape}

In the interest of conciseness and to avoid repeating material that has already been covered in the literature, the discussion on variational integrators has been substantially reduced.
As a result, this table has been removed from the manuscript.

Figure 2 - x-axis has no label but I think it's nondimensional time, please indicate.  Label all axes even if they are nondimensional units.

\item 
    \begin{itshape}
I recommend removing appendix B.  Giving the equations and stating that the Matrix inversion is from Gauss-Jordan elimination is probably sufficient, unless there is some compelling algebraic manipulation that is very unique to the problem.
\end{itshape}

Appendix B has been removed from the revised manuscript.

\item 
    \begin{itshape}
In line 14 (page 12), you define the state transition matrix over one orbital period, and then on line 18 you start referring to it as the monodromy matrix.  I understand what you're doing, but you may want to make the definition more explicit.  Some of this section can be substantially reduced and other source material could be referenced anyway.
\end{itshape}

The discussion about the invariant manifolds and \Poincare sections has been reduced.
This discussion has been removed from the revised manuscript and replaced with a general summary and some relevant citations.

### 2017-07-07 Response.tex

Figure 2.4
Lew, A., Marsden, J., Ortiz, M., West, M., "An Overview of Variational Integrators," FINITE ELEMENT METHODS: 1970's AND BEYOND L.P. Franca (Ed.) c CIMNE, Barcelona, Spain 2003.

Online copy:
http://lagrange.mechse.illinois.edu/pubs/LeMaOrWe2004/LeMaOrWe2004.pdf
\end{itshape}

The comparison plot has been modified to follow the precedent in the literature.
The plot shows the deviation of the Jacobi constant as the time step is varied. 
In addition, an additional plot has been added, and some related discussion which shows the decrease in computation cost through the use of the variational integrator.

\item 
    \begin{itshape}
It may also be generally worth acknowledging the benefits of Taylor's method within the context of variational integrators:
Schmitt, J., Shingel, T., Leok, M., "LAGRANGIAN AND HAMILTONIAN TAYLOR VARIATIONAL INTEGRATORS," eprint arXiv:1703.06599, 2017

Online copy: https://arxiv.org/pdf/1703.06599.pdf 
\end{itshape}

This manuscript is not really focused on variational integrators but rather the design of transfer trajectories using reachability sets.
Interested readers can find a wide variety of literature on the relative benefits of all manner of variational integrators in papers focused on that topic specifically.

\item 
    \begin{itshape}
"Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques".  I think you need to add "in which energy conservation is mandatory for characterizing the solution space."
\end{itshape}

The additional clarification has been added to the manuscript.

### 2017-07-07 manuscript.tex
Add more detail on variational integrators
Reviewer 1 Comment 20, 21

Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques.
\section{Optimal Control Formulation for Reachability Set}
% introduce method
In this section, an optimal control formulation is presented to determine and design transfers within the three-body problem.
The application of variational integrators to optimal control problems is referred to as computational geometric optimal control.
Our formulation is based on the concept of the reachability set on a \Poincare section.
This method allows one to easily determine potential transfer opportunities by finding set intersections on a lower dimensional space and greatly reduces the design process.
The addition of continuous low thrust propulsion extends the control free design process developed previously and allows for a greater range of potential transfers with a reduced time of flight.

>>>>

Over long simulation horizons or with the addition of small control inputs this poor energy behavior limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.

\section{Optimal Control Formulation for Reachability Set}
% introduce method
In this section, an optimal control formulation is presented to determine and design transfers within the three-body problem.
The application of variational integrators to optimal control problems is referred to as computational geometric optimal control.
Our formulation is based on the concept of the reachability set on a \Poincare section.
This method allows one to easily determine potential transfer opportunities by finding set intersections on a lower dimensional space and greatly reduces the design process.
The addition of continuous low thrust propulsion extends the control free design process developed previously and allows for a greater range of potential transfers with a reduced time of flight.
